What words ought to exist? \\ {\normalsize Coining with coinduction}

### 2011-03-27 wishlist/paper.tex
wrote about backformation

\begin{abstract}
what do I put here
\end{abstract}

@@

Since the best research is intensely navel-gazing, I started by 
analyzing a corpus of my own writing, specifically my Facebook status
updates since March~2006. There were 1,386 status updates containing
such jems as ``Tom Murphy VII thinks mathfrak is straight ballin''
and ``Tom Murphy VII global L\_50 reused for unused\_36166!!''. The
most likely words with $n=4$:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
.252\% &  pittsburgh   & .097\% &  can't    \\   
.209\% &  steelers     & .083\% &  i'm      \\   
.209\% &  sfo          & .083\% &  icfp     \\   
.195\% &  it's         & .083\% &  app      \\   
.125\% &  bdl          & .069\% &  x        \\   
.111\% &  sigbovik     & .069\% &  drunj    \\   
.109\% &  facebook     & .069\% &  g        \\   
.097\% &  mic          & .061\% &  ther     \\   
.097\% &  s            & .055\% &  doesn't  \\
\end{tabular}
\end{center}

This is the worst. Not only does it contain loads of one-letter words that
we have already determined are verboten\footnote{Note that since $n=4$ these
words have to actually appear in status updates to have nonzero probability for
this list. ``g'' is explained by frequent occurrences of ``e.g.'', for example.},
but the rest are just non-words that I tend to use like the names of cities,
prestigious conferences, or IATA airport codes.

Wikipedia favorites: smally (.000287),
websity (.000156),
stude (.000156),
chool (.000124),
fontry (.000120),
undex (.000102),
octory (.0000990),
coibot (.0000960),
footnot (.0000837),
reporth (.0000518),
delection (.0000484),
grounty (.0000459),
betweek (.0000437),
fination (.0000431),
manuary (.0000388),
whicle (.0000360),
stategory (.0000262),

@@

The empty string??
% TODO: the empty string?

@@

\begin{verbatim}
Best fixes...
1.000000000 -zzyingly    1/1
1.000000000 -zzying      1/1
1.000000000 -zzuolanas   1/1
1.000000000 -zzuolana    1/1
1.000000000 -zzotints    1/1
1.000000000 -zzotintos   1/1
1.000000000 -zzotinto    1/1
1.000000000 -zzotinting  1/1
1.000000000 -zzotinters  1/1
1.000000000 -zzotinter   1/1
1.000000000 -zzotinted   1/1
1.000000000 -zzotint     1/1
1.000000000 -zzonites    1/1
1.000000000 -zzonite     1/1
1.000000000 -zzoners     1/1
1.000000000 -zzoner      1/1
1.000000000 -zzolans     1/1
1.000000000 -zzolanic    1/1
1.000000000 -zzolan      1/1
1.000000000 -zzmen       1/1
1.000000000 -zzman       1/1
1.000000000 -zzlike      1/1
1.000000000 -zzini       1/1
1.000000000 -zzicatos    1/1
1.000000000 -zzicato     1/1
1.000000000 -zzicati     1/1
1.000000000 -zzian       1/1
\end{verbatim}
1/1 is not really much support.

Ones that are only seen n/1 time never apply (because ...). Filtered:

(actually we should probably be using binomial likelihood)

\begin{verbatim}
Best fixes...
1.000000000 -zzazzes     3/3
1.000000000 -zzazz       3/3
1.000000000 -zzans       3/3
1.000000000 -zzanim      2/2
1.000000000 -zzan        3/3
1.000000000 -zygotic     3/3
1.000000000 -zygotes     2/2
1.000000000 -zygote      2/2
1.000000000 -zygosity    2/2
1.000000000 -zygosis     2/2
1.000000000 -zpas        2/2
1.000000000 -zpahs       2/2
1.000000000 -zpah        2/2
1.000000000 -zpa         2/2
1.000000000 -zooks       2/2
1.000000000 -zonites     3/3
1.000000000 -zonite      3/3
1.000000000 -zonians     2/2
1.000000000 -zonian      2/2
1.000000000 -zists       2/2
1.000000000 -zist        2/2
\end{verbatim}

Not much better!

Only ones with denominator 10+ times.
\begin{verbatim}
Best fixes...
1.000000000 -wrought    10/10
1.000000000 -writings   12/12
1.000000000 -wraps      10/10
1.000000000 -wrap       11/11
1.000000000 -worms      69/69
1.000000000 -worm       69/69
1.000000000 -working    21/21
1.000000000 -worked     14/14
1.000000000 -winds      17/17
1.000000000 -wind       18/18
1.000000000 -wide       12/12
1.000000000 -weights    19/19
1.000000000 -weight     22/22
1.000000000 -waxes      12/12
1.000000000 -wax        12/12
1.000000000 -waters     37/37
1.000000000 -water      41/41
1.000000000 -walking    11/11
1.000000000 -walkers    17/17
1.000000000 -walker     17/17
1.000000000 -valuing    10/10
1.000000000 -valued     11/11
1.000000000 -towns      11/11
1.000000000 -tight      15/15
1.000000000 -thought    12/12
1.000000000 -thinks     11/11
\end{verbatim}

Much better! These are still sort of degenerate and won't help
us find new words because they ALWAYS apply!

Instead anything with p strictly less than 1.
\begin{verbatim}
Best fixes...
0.985714286 -makers     69/70
0.985714286 -maker      69/70
0.983606557 -wood       120/122
0.983471074 -woods      119/121
0.982758621 -down       57/58
0.982658960 -works      170/173
0.981818182 -houses     108/110
0.981818182 -house      108/110
0.981132075 kilo-       52/53
0.980752406 -less       1121/1143
0.980743395 over-       2190/2233
0.980000000 -books      49/50
0.980000000 -book       49/50
0.979591837 -proof      48/49
0.979310345 -lessnesses 142/145
0.979069767 -ships      421/430
0.978723404 -lessness   184/188
0.978723404 -board      138/141
0.978494624 -woman      91/93
0.978021978 -women      89/91
0.977528090 -ship       435/445
0.977272727 -manship    43/44
0.976744186 -weeds      84/86
0.976470588 after-      83/85
0.976190476 -manships   41/42
0.976190476 -making     41/42
0.976190476 -craft      41/42
0.976190476 -boats      41/42
0.976190476 -boat       41/42
\end{verbatim}

Wow, now we're talking!

@@

From this we can conclude that 16\% of people wish {\it nurm} was a word, and so on.
These words did not come with definitions, except for {\it gruntle}, which Jessica
gives as ``the opposite of distgruntle''. This is actually already a word, but
it was the inspiration for Section~\ref{sec:backformation}. {\it etsy} is the name
of a popular on-line crafts community so I don't know why Rob would suggest that.
The meaning of {\it wafflucinations} is clear from morphological analysis.

>>>>

\begin{abstract}
This paper is an earnest attempt to answer the following question
scientifically: What words ought to exist?
\end{abstract}

@@

{\bf Facebook.}\ Since the best research is intensely navel-gazing, I started by 
analyzing a corpus of my own writing, specifically my Facebook status
updates since March~2006. There were 1,386 status updates containing
such jems as ``Tom Murphy VII thinks mathfrak is straight ballin''
and ``Tom Murphy VII global L\_50 reused for unused\_36166!!''. The
most likely words with $n=4$:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{My Facebook status updates, $n=4$.} \\
\hline
.252\% &  pittsburgh   & .097\% &  can't    \\   
.209\% &  steelers     & .083\% &  i'm      \\   
.209\% &  sfo          & .083\% &  icfp     \\   
.195\% &  it's         & .083\% &  app      \\   
.125\% &  bdl          & .069\% &  x        \\   
.111\% &  sigbovik     & .069\% &  drunj    \\   
.109\% &  facebook     & .069\% &  g        \\   
.097\% &  mic          & .061\% &  ther     \\   
.097\% &  s            & .055\% &  doesn't  \\
\end{tabular}
\end{center}

This is the worst. Not only does it contain loads of one-letter words that
we have already determined are verboten,\!\footnote{Note that since $n=4$ these
words have to actually appear in status updates to have nonzero probability for
this list. ``g'' is explained by frequent occurrences of ``e.g.'', for example.}
but the rest are just non-words that I tend to use like the names of cities,
prestigious conferences, or IATA airport codes. The main problem is that there
is simply not enough data from which to generalize.

{\bf Wikipedia.}\ I tried again, but with Wikipedia, using a snapshot
of the English site from June~2009. This is 23 gigabytes of data, most
of it expository text composed by native speakers, plus bathroom humor
vandalism. The list produced by this analysis is much better, though it
contains artifacts from non-English Wiki language used in articles. The
unabridged list appears in the appendix; my hand-selected favorites:

\begin{center}
\begin{tabular}{rl@{\quad\quad}rl}
\multicolumn{4}{c}{English Wikipedia, $n=3$.}     \\
\hline
.0287\% & smally      &     .00518\% & reporth    \\
.0156\% & websity     &     .00484\% & delection  \\
.0156\% & stude       &     .00459\% & grounty    \\
.0124\% & chool       &     .00437\% & betweek    \\
.0120\% & fontry      &     .00431\% & fination   \\
.0102\% & undex       &     .00388\% & manuary    \\
.0099\% & octory      &     .00360\% & whicle     \\
.0096\% & coibot      &     .00262\% & stategory  \\
.0084\% & footnot     &                           \\
\end{tabular}
\end{center}

Lots of these could be the names of tech startups or Pok\'emon.

@@

The empty string?? Is that a word? Could it be? Dude that is blowing my mind.

@@

The lexicon is generative, in the sense that it's possible to make new
words that are generally acceptable, by following rules. Most people
recognize pluralization of nouns by adding {\it --s} (even for novel
words), or adding prefixes like {\it anti--}. We could investigate
words that ought to exist by the application of rules, such as {\it
  examplelikelikelikelikelikelike}, but I see no straightforward way
to justify the relative strength of such words.

A related way for words to enter the lexicon is by backformation. This
is the reverse of the above process: A word like {\it laser}
(initially an initialism) is legal, and then by running the rules of
English backwards, we start to use {\it lase} as a word (the verb
that a laser most frequently applies). In this section, I attempt
to determine formation rules in English (by simple lexical analysis
of the set of legal words) and then run these rules backwards to find
words that seemingly should already exist.

{\bf Prefixes and suffixes.}\ The first order of business is to find
prefixes and suffixes that are usually modular. The kind of thing
we're tring to find are ``anti--'' and ``--ing''; stuff you can often
add to a word to make a related word. The approach is straightforward.
For each word, consider splitting it at each position. For {\it dealing},
we have {\it d/ealing}, {\it de/aling}, etc. For every such split,
take the prefix (e.g. ``de'') and remainder (``aling''); if the remainder
is still a legal word, then the prefix gets one point. {\it aling} is not
a word so no points here for ``de''. We also do the same thing for suffixes
(using the exact same splits, symmetrically). In this case we'll only get
points for ``--ing'' since {\it deal} is a word. Every time a prefix or
suffix appears we test to see if it is being applied modularly, and
the final score is just the fraction of such times. Here are the ones with
the highest scores:

\begin{verbatim}
1.000000000 -zzyingly    1/1
1.000000000 -zzying      1/1
1.000000000 -zzuolanas   1/1
1.000000000 -zzuolana    1/1
1.000000000 -zzotints    1/1
1.000000000 -zzotintos   1/1
1.000000000 -zzotinto    1/1
1.000000000 -zzotinting  1/1
...
\end{verbatim}

Well, it's good to know that 100\% of the time, you can remove
``--zzotinting'' from a word and it will still be a word. But this
inference is supported by just one observation (the word {\it
  mezzotinting}); there are actually hundreds of such unique prefixes
and suffixes. We need a better list.\!\footnote{The right thing to do
  here is probably to use binomial likelihood rather than the
  scale-independent fraction. But simpler approaches produce pretty
  good lists.} Removing the ones that appear just a single time
doesn't really help that much:

\begin{verbatim}
1.000000000 -zzazzes     3/3
1.000000000 -zzazz       3/3
1.000000000 -zzans       3/3
1.000000000 -zzanim      2/2
1.000000000 -zzan        3/3
1.000000000 -zygotic     3/3
\end{verbatim}

Still bad. Let's turn up the juice to prefixes and suffixes that
appear at least 10 times.

\begin{verbatim}
1.000000000 -wrought    10/10
1.000000000 -writings   12/12
1.000000000 -wraps      10/10
1.000000000 -wrap       11/11
1.000000000 -worms      69/69
1.000000000 -worm       69/69
1.000000000 -working    21/21
\end{verbatim}

Much better! But the next step is going to be to try removing these
prefixes and suffixes from words that have them, to find new words.
Since these have modularity of 100\%, we already know that every
time we apply them, the result will already be a word. So they are
useless for our analysis. Here are the most modular prefixes
and suffixes with modularity {\it strictly less than} 1.

\begin{verbatim}
0.985714286 -makers     69/70
0.985714286 -maker      69/70
0.983606557 -wood       120/122
0.983471074 -woods      119/121
0.982758621 -down       57/58
0.982658960 -works      170/173
0.981818182 -houses     108/110
0.981818182 -house      108/110
0.981132075 kilo-       52/53
0.980752406 -less       1121/1143
0.980743395 over-       2190/2233
0.980000000 -books      49/50
0.980000000 -book       49/50
0.979591837 -proof      48/49
0.979310345 -lessnesses 142/145
0.979069767 -ships      421/430
0.978723404 -lessness   184/188
0.978723404 -board      138/141
0.978494624 -woman      91/93
0.978021978 -women      89/91
0.977528090 -ship       435/445
0.977272727 -manship    43/44
0.976744186 -weeds      84/86
0.976470588 after-      83/85
0.976190476 -manships   41/42
0.976190476 -making     41/42
0.976190476 -craft      41/42
0.976190476 -boats      41/42
0.976190476 -boat       41/42
\end{verbatim}

Wow, now we're talking! The single word that cannot have ``--maker''
removed is {\it comaker}, suggesting that {\it co} should be word
(noun: ``What a comaker makes.'').

Given this list, the next step is to identify potential words that
can be backformed by removing prefixes or adding suffixes from existing
words. Such a string can often be found via multiple prefixes and
suffixes. For example, {\it twing} can be formed by removing ``--ing''
from {\it twinging} (a false positive, since the root word is actually
{\it twinge} in this case) as well as by removing the prefix ``lef--'',
which has modularity of 20\% (including splits such as ``lef/tie'').
Maybe not good justification, but {\it twing} is a pretty good word
anyway.

We define the probability of a word as its Markov probability (with
$n=4$, as this seems to produce the best results), times the
probability that at least one of the potential backformation rules
applies.\!\footnote{As above we only allow backformation rules that
  have at least 10 occurrences, to prevent degeneracy.} Here are the
most likely words by backformation:

\begin{verbatim}
word    prob   most likely backformation rules
==============================================
dises   .023%  para- (0.42) fluori- (0.39) 
               melo- (0.35) bran- (0.31)  
tring   .020%  hams- (0.36) scep- (0.35) 
               bows- (0.33) hearts- (0.29)  
disms   .017%  triba- (0.31) drui- (0.30) 
               bar- (0.27) invali- (0.27)  
ching   .017%  day- (0.86) hot- (0.69) 
               star- (0.51) guillo- (0.50)    
sking   .017%  dama- (0.24) imbo- (0.18) 
               fri- (0.18) atta- (0.17)     
cally   .015%  anti- (0.78) specifi- (0.61)
               magnifi- (0.55) phoni-    
pring   .015%  days- (0.67) heads- (0.62)
               outs- (0.54) ups- (0.51)    
\end{verbatim}

I think that this approach shows promise, but there appear to be a few
problems: Many of these ``rules'' can be explained by bad segmentation
(``heads--'' appearing to be modular, for example, is really just
``head--'' plus ``s'' being a common letter.) Second, I believe the
disjunctive probability of any rule applying is too naive for
determining the score. For example, {\it tions} has almost a thousand
different prefixes that could apply to it; the chance of {\it any one}
of them applying is very nearly 1. But this is actually because ``tions''
is just a common way for a word to end. Legitimate root words to which
many good prefixes are applied cannot be easily distinguished from
common suffixes by this symmetric algorithm. More work is called for
here.

@@

From this we can conclude that 16\% of people wish {\it nurm} were a word, and so on.
These words did not come with definitions, except for {\it gruntle}, which Jessica
gives as ``the opposite of distgruntle''. This is actually already a word, but
it was the inspiration for Section~\ref{sec:backformation}. {\it etsy} is the name
of a popular on-line crafts community so I don't know why Rob would suggest that.
The meaning of {\it wafflucinations} is clear from morphological analysis.
